### 2024-03-15 paper/paper.qmd
First render and some tests

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(arrow)
library(downloader)
library(rstanarm)
```

# Introduction

# Data

# Results

```{r}
#| label: fig-total_yearly_death
#| fig-cap: "Visual of the total deaths in Ontairo from 2000 - 2022."
#| eval: true
#| echo: false
#| message: false
#| warning: false

cleaned_data <- read_csv("/Users/liamwall/Ontario-CA-Mortality/data/analysis_data/cleaned_data.csv")

#Total annual death data for both sexes and all causes
total_yearly_death_data <- cleaned_data |>
  filter(characteristics == "Number of deaths",
         cause == "Total, all causes of death [A00-Y89]",
         age == "Age at time of death, all ages",
         sex == "Both sexes") |>
  mutate(number_of_deaths = value) |>
  select(year, age, sex, cause, number_of_deaths) |>
  ggplot(aes(x = year, y = number_of_deaths)) +
  geom_point()

total_yearly_death_data
```

# Discussion

# Appendix

## Additional data details

## Model details {#sec-model-details}

## Posterior predictive check

# References